Changes in Mobility and Site Occupation during the Late Pleistocene in Korea

### 2020-01-09 analysis/paper/paper.Rmd
Done with China, move to Japan

The Shuidonggou site in northern China is a typical site for showing the transition from cobble to blade tool industries in the Late Pleistocene of Northeast Asia [@Brantingham_Perreault_2010; @Gao_Zhang_Yang_Shen_Wu_2010; @pei2012shuidonggou]. Shuidonggou Localities SDG 2, 7-12 have blades and bladelets made of higher quality or exotic raw materials in addition to more than 80 polished ostrich egg-shell beads (SDG 2), bone needles and an awl (SDG 12) [@pei2012shuidonggou]. The blade compoenent is claimed to have come from Siberia and/or Mongolia at 41 ka, and at 10.5 ka  microblade technology appears but the origin is uncertain. <!-- expand on ages of all the things, comment about the cause of the new things at this site, check, https://link.springer.com/article/10.1007/s10963-019-09129-w  what is the debate? -->

[@yi2014significance; @gao2014discovery; @zhang2016bone]

The origins of Late Pleistocene technological innovations in Japan is debated due to the uncertainty about the existence of people in the Archipelago before the appearance of new tools (citations). New technologies are represented by systematic procurement and reduction of lithic raw materials, trap-pit hunting methods, watercraft, bow-and-arrow technology, and standardization of lithic tools. <!-- need some ages for these types of evidence --> These have been interpreted as evidence of new foraging strategies possibly driven by increased population and ecological changes [@Yamaoka_2012; @bae2017origin; @Nakazawa_Bae_2018; @morisaki2019early]. <!-- need to expand, mention some key site, blades appear when? what is the debate? -->

>>>>

The Shuidonggou site in northern China is a typical site for showing the transition from cobble to blade tool industries in the Late Pleistocene of Northeast Asia from the Late Pleistocene to Middle Holocene [@Brantingham_Perreault_2010; @Gao_Zhang_Yang_Shen_Wu_2010; @pei2012shuidonggou]. Shuidonggou Localities SDG 2, which dated around c.32,000 years, contains aunal specimens,ostrich eggshell beads,and stone artefacts including blades, retouched flakes, cores, and debries along with well-preserved hearths. In Locality 9 dated to c.29,000years, there is a small scatter of stone artefacts consists of blades, Levallois flakes, cores and other retouched flakes. <!-- Levallois! I didn't pay attentioned --> The recent excavtion of Locality 12 shows that its archaeological layer <!--I shouldn't use 'layer', but I don't know what to say shortly-->, dated to 13,078–13,296 cal BP, includes ground stone and more than 30,000 microlithics made out of a variety of raw materials such as fine-grained and siliceous rocks. In addition, more than 10,000 animal fossils and bone tools including a tool for fishing nets, an awl, and two needles were excavated [@pei2012shuidonggou; @gao2014discovery; @zhang2016bone].<!-- they use OSL data that has wise range, or calibrated years --> Based on the archaeological records, people argue that a microblade technology appeared in China after 29ka BP related to mobility change along with colder climate towards LGM and changed game as surviving strategy [@yi2014significance]. The accumulated studies on Shuidonggou site show that the blade compoenent is claimed to have come from Siberia and/or Mongolia at 41 ka (SDG Locality1 and 9), advanced core and flake tools are likely to be locally developed at 33ka (2 and 8), and at 10.8 ka microblade technology appears but the origin is uncertain (12) [@yi2014significance; @li2019history].

<!-- the last sentense is a bit long-->
<!-- "advanced core and flake tools means intentionally prepared platforms and working surfaces, selective raw material usage (using chert rather tahn river pebble), bipolar reduction, and generating small elongated blanks as results" -->

The origins of Late Pleistocene technological innovations in Japan is debated due to the uncertainty about the existence of people in the Archipelago before the appearance of new tools

[@izuho201421; ]

. New technologies are represented by systematic procurement and reduction of lithic raw materials, trap-pit hunting methods, watercraft, bow-and-arrow technology, and standardization of lithic tools [@bae2017origin]. <!-- need some ages for these types of evidence --> These have been interpreted as evidence of new foraging strategies possibly driven by increased population and ecological changes [@Yamaoka_2012; @bae2017origin; @Nakazawa_Bae_2018; @morisaki2019early]. <!-- need to expand, mention some key site, blades appear when? what is the debate? -->
